### 2024-12-10 Exam Assignments/Chapters/03/03.tex
Week 09

~\\
The outer most loop cannot be parallelized, as their iterations are all strictly dependent on the previous one.
This is why the implicit barrier in line 13 is neccessary.
It also shouldn't be that big of a bottleneck, as the threads should reach it almost simultaneously.

~\\
\subsubsection{Optimizing Random Shooting}
When trying to approximate $\pi$ by the method of random shooting we use random numbers.
They, however, don't need to be truly random, they just have to simulate it well enough.
One way to do this simulation of randomness is the rnd function like it is defined in the book (or see the code below).
It calculates a random number quite efficiently.
Another benfit is that we do not need multiple instances of random number generators, and we also don't run the risk of misusing those by calling their methods in different threads.
The code is shown in the \hyperref[fig:calculate_pi_v2]{figure} below.
It runs very quickly and calculates $\pi$ very precisely.
The running time on my particular machine was 0.0313137 s, in comparision to the previous version with a running time of 0.117771 s.

>>>>

~\\
The outer most loop cannot be parallelized, as their iterations are all strictly dependent on the previous one.
This is why the implicit barrier in line 13 is necessary.
It also shouldn't be that big of a bottleneck, as the threads should reach it almost simultaneously.

~\\
\subsubsection{Optimizing Random Shooting}
When trying to approximate $\pi$ by the method of random shooting we use random numbers.
They, however, don't need to be truly random, they just have to simulate it well enough.
One way to do this simulation of randomness is the rnd function like it is defined in the book (or see the code below).
It calculates a random number quite efficiently.
Another benfit is that we do not need multiple instances of random number generators, and we also don't run the risk of misusing those by calling their methods in different threads.
The code is shown in the \hyperref[fig:calculate_pi_v2]{figure} below.
It runs very quickly and calculates $\pi$ very precisely.
The running time on my particular machine was 0.0313137 s, in comparison to the previous version with a running time of 0.117771 s.

### 2024-12-10 Exam Assignments/Chapters/07/07.tex
Week 09

\begin{document}
\newcommand{\drawcells}[3]{%
    \begin{tikzpicture}
        \foreach \j in {1,...,#1} {
            % Draw each cell
            \node[draw, fill=#3!20, minimum width = \linewidth/#1, minimum height=0.5cm, align=center] (cell\j) at (\j*\linewidth/#1, 0) {#2};
        }
    \end{tikzpicture}}
\section{Guided Vectorization and Data Types for Vector Intrinsics}
\subsection{Vectorization Clauses}
\begin{description}
    \item[aligned] ~\\ This clause tells the compiler that every element of an array $a$ is $n$-bit aligned if one uses \texttt{\#pragma opm simd aligned(a, n)}. The compiler employs this information to effectively vectorize the code section.
    \bigskip
    \item[safelen] ~\\ Imagine we had the following code:
    \begin{lstlisting}
a[0] = b[0];
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}
    We have a data dependency, but only every second iteration: iteration $i$, where $i$ is odd, depends on $i-1$. Iteration $i$ and $i+1$ are completely independent, however. Hence, we can vectorize the loop wath a vector length of $2$. To specify this, we use the \texttt{safelen} clause like so:
    \begin{lstlisting}
a[0] = b[0];
#pragma omp simd safelen(2)
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}   
    \bigskip
    \item[reduction] ~\\ This clause has the same functionality when paired with \texttt{simd} as when paired with \texttt{paraller for}: It will reduce a variable by an associative operation. But instead of reducing it over multiple threads, we use vectorization for reducing. I imagine that the reduced variable itself will be vectorized, so that the in loop we can employ vector operations, and at the end the elements in the resulting vector will get reduced.
\end{description}

@@

    This is a typical example of branching inside a loop, where assignments have the following form:
    \[
        \texttt{a[i] = expr\_1 if cond else expr\_2}
    \]
    The idea is that we can vectorize this using vector istructions by:
    \[
        \texttt{a[i] = (expr\_1 $\land$ cond) $\lor$ (expr\_2 $\land$ $\lnot$cond)}
    \]
    Where \texttt{cond} is evaluated to \texttt{111\dots111} if true, and to \texttt{000\dots000} if false. The expressions should be easily vectorizable themselves, such that ultimately the entire statement can be vectorized efficiently.

@@

    \bigskip
    \item[Matrix Transposition] Vector instructions can be fully utilized if the data is stored sequentially in memory. Hence, accessing rows of matrices can vectorized. But I we want to work on the columns of the matrix, it might be beneficial to transpose the matrix. This transposing can be vectorized as well by employing the so called \em unpack \em operation, which interlieves two operands word-wise, as can be seen in the picture:

>>>>

\begin{document}
\newcommand{\drawcells}[3]{%
    \begin{tikzpicture}
        \foreach \j in {1,...,#1} {
            % Draw each cell
            \node[draw, fill=#3!20, minimum width = \linewidth/#1, minimum height=0.5cm, align=center] (cell\j) at (\j*\linewidth/#1, 0) {#2};
        }
    \end{tikzpicture}}
\section{Guided Vectorization and Data Types for Vector Intrinsics}
\subsection{Vectorization Clauses}
\begin{description}
    \item[aligned] ~\\ This clause tells the compiler that every element of an array $a$ is $n$-bit aligned if one uses \texttt{\#pragma opm simd aligned(a, n)}. The compiler employs this information to effectively vectorize the code section.
    \bigskip
    \item[safelen] ~\\ Imagine we had the following code:
    \begin{lstlisting}
a[0] = b[0];
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}
    We have a data dependency, but only every second iteration: iteration $i$, where $i$ is odd, depends on $i-1$. Iteration $i$ and $i+1$ are completely independent, however. Hence, we can vectorize the loop with a vector length of $2$. To specify this, we use the \texttt{safelen} clause like so:
    \begin{lstlisting}
a[0] = b[0];
#pragma omp simd safelen(2)
for (int i = 1; i < n; i++) {
    a[i] = a[i-1] * (i % 2) + b[i];
}
    \end{lstlisting}   
    \bigskip
    \item[reduction] ~\\ This clause has the same functionality when paired with \texttt{simd} as when paired with \texttt{paraller for}: It will reduce a variable by an associative operation. But instead of reducing it over multiple threads, we use vectorization for reducing. I imagine that the reduced variable itself will be vectorized, so that the in loop we can employ vector operations, and at the end the elements in the resulting vector will get reduced.
\end{description}

@@

    This is a typical example of branching inside a loop, where assignments have the following form:
    \[
        \texttt{a[i] = expr\_1 if cond else expr\_2}
    \]
    The idea is that we can vectorize this using vector instructions by:
    \[
        \texttt{a[i] = (expr\_1 $\land$ cond) $\lor$ (expr\_2 $\land$ $\lnot$cond)}
    \]
    Where \texttt{cond} is evaluated to \texttt{111\dots111} if true, and to \texttt{000\dots000} if false. The expressions should be easily vectorizable themselves, such that ultimately the entire statement can be vectorized efficiently.

@@

    \bigskip
    \item[Matrix Transposition] Vector instructions can be fully utilized if the data is stored sequentially in memory. Hence, accessing rows of matrices can vectorized. But I we want to work on the columns of the matrix, it might be beneficial to transpose the matrix. This transposing can be vectorized as well by employing the so called \em unpack \em operation, which interleaves two operands word-wise, as can be seen in the picture:

### 2024-12-10 Exam Assignments/Chapters/08/08.tex
Week 09

~\\
On the other hand, if we write independent code which also uses many repeated operations, throughput will be key, as it determines how many clock cycles we approximately need to finish all calculations.
%  \\ ($no\_clock\_cycles \approx \frac{no\_operations}{throughput\_of\_operation}$ (Note that the number of operations does not directly correlate with the number of instructions in code, as certain code instructions need more cpu instructions)).
In this case, the latency has a neglectable influence on the perfoncmance, especially if the number of operations grows bigger.

@@

\begin{description}
    \item[Vectorized Linear Probing] The keys should be integer values, so that they can fit comfortably into the vector registers. Linear probing starts by checking the index specified by the hash value of the current key first, and then keeps going one step at a time until we found our key. The "key" idea is to simply load multiple conscutive keys into one vector register, and to compare them all against the search key. If one comparision is successful, we can calculate the index and return tha value at that index. Otherwise, we keep probing with the next set of keys.

>>>>

~\\
On the other hand, if we write independent code which also uses many repeated operations, throughput will be key, as it determines how many clock cycles we approximately need to finish all calculations.
%  \\ ($no\_clock\_cycles \approx \frac{no\_operations}{throughput\_of\_operation}$ (Note that the number of operations does not directly correlate with the number of instructions in code, as certain code instructions need more cpu instructions)).
In this case, the latency has a neglectable influence on the performance, especially if the number of operations grows bigger.

@@

\begin{description}
    \item[Vectorized Linear Probing] The keys should be integer values, so that they can fit comfortably into the vector registers. Linear probing starts by checking the index specified by the hash value of the current key first, and then keeps going one step at a time until we found our key. The "key" idea is to simply load multiple consecutive keys into one vector register, and to compare them all against the search key. If one comparison is successful, we can calculate the index and return tha value at that index. Otherwise, we keep probing with the next set of keys.
